Dataflow-Guided Filtering for Efficient and Adjustable Run-Time Monitoring

### 2014-05-30 optimizations.tex
Add figure for policy trade off.

The instruction-grained run-time monitoring schemes we focus on typically
manage a set of metadata information. This metadata information is used to
verify security or reliability probabilities. For example, for array bounds
check, the monitoring core stores metadata describing the base and bound
addresses for array pointers. This metadata is then checked on load and store
instructions. However, if the metadata for an event is null, then there is no need to perform the check since this represents a non-array pointer value. By
identifying and filtering out these null monitoring events, we can reduce the overheads of
monitoring by reducing the amount of work the monitoring core must do.

In order to filter out null monitoring events, we need an efficient method to
keep track of which events are null. In addition, when we filter out a null
monitoring event, we must ensure that any metadata updates that were skipped
are handled properly.  Thus, when we filter out one of
these events, we must ensure that the destination operand is also marked as null. 
In other words, we need to be able to track not only
metadata which is initially null, but also the dataflow of null metadata.

>>>>

The instruction-grained run-time monitoring schemes we focus on typically
manage a set of metadata information. This metadata information is used to
verify security or reliability properties. For example, for array bounds
check, the monitoring core stores metadata describing the base and bound
addresses for array pointers. This metadata is then checked on load and store
instructions. However, if the metadata for an event is null, then there is no
need to perform the check since this represents a non-array pointer value. By
identifying and filtering out these null monitoring events, we can reduce the
overheads of
monitoring by reducing the amount of work the monitoring core must do.

In order to filter out null monitoring events, we need an efficient method to
keep track of which events are null. In addition, when we filter out a null
monitoring event, we must ensure that any metadata propagations that were
skipped are handled properly. If this is not done, the monitoring core may
operate on stale metadata. Thus, when we filter out a null monitoring event,
we must ensure that the destination metadata is also marked as null. 
In other words, we need to be able to track not only
metadata which is initially null, but also the dataflow of null metadata.

### 2014-05-30 policies.tex
Add figure for policy trade off.

% Slack
\begin{figure}
  \begin{center}
    \includegraphics[width=\columnwidth]{figs/slack.pdf}
    \vspace{-0.3in}
    \caption{Slack and its effect on monitoring over time.}
    \label{fig:drop.slack}
    \vspace{-0.1in}
  \end{center}
\end{figure}

% Slack tracking module
\begin{figure}
  \begin{center}
    \includegraphics[width=\columnwidth]{figs/stm.pdf}
    \vspace{-0.2in}
    \caption{Slack tracking and drop decision hardware.}
    \label{fig:drop.stm}
    \vspace{-0.1in}
  \end{center}
\end{figure}

@@

% Instead, we develop a model to estimate the execution time overhead budget in a
% gradual manner and without the need to modify the main program. Specifically,
We specify the overhead budget as a percentage of the main program's execution
cycles without monitoring. We define \emph{slack} as the number of cycles of
monitoring overhead that can be incurred while staying within the budget
target. Slack is essentially the difference between the actual overheads seen
and the budget specified. Slack is generated as the main program runs and consumed as monitoring overheads occur.  For
example, if no monitoring overheads occur during 1000 cycles of the main
program's execution and the designer sets a 20\% overhead target, then the
slack that is built up during this period is 200 cycles. If the main core is
then stalled for 50 cycles due to monitoring, then the remaining slack is 150
cycles. 
In addition to this accumulated slack, a small amount of initial slack
can be given in order for monitoring to be performed at the start of a program.
Figure~\ref{fig:drop.slack} shows an example of how slack can change over time.
If the slack falls below zero (i.e., the overhead budget is exceeded), then
monitoring events should be dropped.

Figure~\ref{fig:drop.stm} shows a hardware slack tracking module for
keeping track of slack. The slack tracking module uses a counter that increments on every $k$-th
cycle of the main core. This $k$ can be calculated by taking the reciprocal of
the target budget. For example, if the target budget is 20\%, then the counter
increments on every 5th clock cycle. The value of this counter is the
accumulated slack. Whenever the main core is stalled due to the monitor, slack
is decremented.

@@

% Varying slack's impact on coverage for UMC
\begin{figure}
  \begin{center}
    \includegraphics[width=\columnwidth]{figs/dataflow_graph.pdf}
    \vspace{-0.3in}
    \caption{Example dependence graph for metadata. Square nodes represent
    events where checks are performed.} 
    \label{fig:drop.dataflow_graph}
    \vspace{-0.1in}
  \end{center}
\end{figure}

The simplest policy for when to drop monitoring events is to drop all monitoring events as
long as slack is less than or equal to zero.  However, this can result in
wasted work. For example, consider the metadata dependence graph shown in
Figure~\ref{fig:drop.dataflow_graph}. Here, an edge from node {\tt A} to node
{\tt B} represents that if event {\tt A} is dropped, then due to its
invalidated metadata, it will cause event {\tt B} to be filtered. Square
nodes indicate events where monitoring checks are performed. In the
example, suppose that event {\tt E} is meant to perform a check operation but is dropped.
In this situation, the
monitoring operations that were done for events {\tt C} and {\tt D} were wasted
since their results were not used for any monitoring checks.

\begin{figure}
  \begin{center}
  \subfloat[unrestricted dropping]{
    \includegraphics[width=\columnwidth]{figs/all_drop.pdf}
    \label{fig:drop.all_drop}
  }
  \vspace{-0.1in}
  \\
  \subfloat[source-only dropping]{
    \includegraphics[width=\columnwidth]{figs/source_drop.pdf}
    \label{fig:drop.source_drop}
  }
  \vspace{-0.1in}
  \\
  \subfloat[split-point dropping]{
    \includegraphics[width=\columnwidth]{figs/min_drop.pdf}
    \label{fig:drop.min_drop}
  }
  \end{center}
  \vspace{-0.2in}
  \caption{Comparison of dropping policies. Blue (dark) nodes indicate where
  dropping decisions are made.}
  \label{fig:drop.policies}
\end{figure}

One method to eliminate this wasted work is to only make dropping decisions at
the root of these metadata flows. That is, we will decide to either monitor or
not monitor an entire metadata flow. We refer to this dropping decision policy
as \emph{source-only dropping} and we refer to the previous policy of 
dropping any event as \emph{unrestricted dropping}.
Figures~\ref{fig:drop.all_drop} and \ref{fig:drop.source_drop} show a
comparison of where dropping decisions are made for these two policies. Source
dropping will
result in no wasted work and thus better coverage. However, because of the coarser-grained decision, it
may be more difficult to closely match overheads. Source dropping is expected
to work well when there are a large number of independent metadata flows.

Another possible method is to make the dropping decisions somewhere in between
source-only dropping and unrestricted dropping. From the graph in 
Figure~\ref{fig:drop.dataflow_graph}, we can see that if the dropping
decision is made at event {\tt C} then we can skip event {\tt E} with no wasted
work. If we instead choose to perform monitoring for event {\tt C}, then we want to perform all
monitoring on that flow. Similarly, event {\tt F} creates a decision point for
the bottom flow that will result in no wasted work and dropping event {\tt A}
allows the entire shown metadata flow to be dropped with no wasted work. If the
program can be
analyzed or profiled to identify these instructions which, if dropped, lead to
no wasted work, then this information can be used at run-time to minimize the
amount of wasted work. We refer to this policy as \emph{sub-flow dropping}.
Figure~\ref{fig:drop.min_drop} shows the points were dropping decisions are
made for this policy.
Min-event dropping is a coarser-grained decision than unrestricted dropping, but
finer-grained that source dropping. Thus, its ability to match overheads is
expected to sit between these other two policies. Similarly, there still exists
edge cases where wasted work can occur, but typically we expect much less
wasted work than unrestricted 
dropping. Thus, we expect the coverage achieved by split-point dropping for a
certain overhead to fall between source and unrestricted dropping. These three
dropping policies create a trade-off space between coverage achieved and
ability to match an overhead target.

>>>>

% Slack
\begin{figure}
  \begin{center}
    \includegraphics[width=\columnwidth]{figs/slack.pdf}
    \vspace{-0.3in}
    \caption{Slack and its effect on monitoring over time.}
    \label{fig:policies.slack}
    \vspace{-0.1in}
  \end{center}
\end{figure}

% Slack tracking module
\begin{figure}
  \begin{center}
    \includegraphics[width=\columnwidth]{figs/stm.pdf}
    \vspace{-0.2in}
    \caption{Slack tracking and drop decision hardware.}
    \label{fig:policies.stm}
    \vspace{-0.1in}
  \end{center}
\end{figure}

@@

% Instead, we develop a model to estimate the execution time overhead budget in a
% gradual manner and without the need to modify the main program. Specifically,
We specify the overhead budget as a percentage of the main program's execution
cycles without monitoring. We define \emph{slack} as the number of cycles of
monitoring overhead that can be incurred while staying within the budget
target. Slack is essentially the difference between the actual overheads seen
and the budget specified. Slack is generated as the main program runs and consumed as monitoring overheads occur.  For
example, if no monitoring overheads occur during 1000 cycles of the main
program's execution and the designer sets a 20\% overhead target, then the
slack that is built up during this period is 200 cycles. If the main core is
then stalled for 50 cycles due to monitoring, then the remaining slack is 150
cycles. 
In addition to this accumulated slack, a small amount of initial slack
can be given in order for monitoring to be performed at the start of a program.
Figure~\ref{fig:policies.slack} shows an example of how slack can change over time.
If the slack falls below zero (i.e., the overhead budget is exceeded), then
monitoring events should be dropped.

Figure~\ref{fig:policies.stm} shows a hardware slack tracking module for
keeping track of slack. The slack tracking module uses a counter that increments on every $k$-th
cycle of the main core. This $k$ can be calculated by taking the reciprocal of
the target budget. For example, if the target budget is 20\%, then the counter
increments on every 5th clock cycle. The value of this counter is the
accumulated slack. Whenever the main core is stalled due to the monitor, slack
is decremented.

@@

% Varying slack's impact on coverage for UMC
\begin{figure}
  \begin{center}
    \includegraphics[width=\columnwidth]{figs/dataflow_graph.pdf}
    \vspace{-0.3in}
    \caption{Example dependence graph for metadata. Square nodes represent
    events where checks are performed.} 
    \label{fig:policies.dataflow_graph}
    \vspace{-0.1in}
  \end{center}
\end{figure}

The simplest policy for when to drop monitoring events is to drop all monitoring events as
long as slack is less than or equal to zero.  However, this can result in
wasted work. For example, consider the metadata dependence graph shown in
Figure~\ref{fig:policies.dataflow_graph}. Here, an edge from node {\tt A} to node
{\tt B} represents that if event {\tt A} is dropped, then due to its
invalidated metadata, it will cause event {\tt B} to be filtered. Square
nodes indicate events where monitoring checks are performed. In the
example, suppose that event {\tt E} is meant to perform a check operation but is dropped.
In this situation, the
monitoring operations that were done for events {\tt C} and {\tt D} were wasted
since their results were not used for any monitoring checks.

\begin{figure}
  \begin{center}
  \subfloat[unrestricted dropping]{
    \includegraphics[width=\columnwidth]{figs/all_drop.pdf}
    \label{fig:policies.all_drop}
  }
  \vspace{-0.1in}
  \\
  \subfloat[source-only dropping]{
    \includegraphics[width=\columnwidth]{figs/source_drop.pdf}
    \label{fig:policies.source_drop}
  }
  \vspace{-0.1in}
  \\
  \subfloat[split-point dropping]{
    \includegraphics[width=\columnwidth]{figs/min_drop.pdf}
    \label{fig:policies.min_drop}
  }
  \end{center}
  \vspace{-0.2in}
  \caption{Comparison of dropping policies. Blue (dark) nodes indicate where
  dropping decisions are made.}
  \label{fig:policies.policies}
\end{figure}

One method to eliminate this wasted work is to only make dropping decisions at
the root of these metadata flows. That is, we will decide to either monitor or
not monitor an entire metadata flow. We refer to this dropping decision policy
as \emph{source-only dropping} and we refer to the previous policy of 
dropping any event as \emph{unrestricted dropping}.
Figures~\ref{fig:policies.all_drop} and \ref{fig:policies.source_drop} show a
comparison of where dropping decisions are made for these two policies. Source
dropping will
result in no wasted work and thus better coverage. However, because of the coarser-grained decision, it
may be more difficult to closely match overheads. Source dropping is expected
to work well when there are a large number of independent metadata flows.

% Slack tracking module
\begin{figure}
  \begin{center}
    \includegraphics[width=\columnwidth]{figs/policy_trade_off.pdf}
    \vspace{-0.3in}
    \caption{Slack tracking and drop decision hardware.}
    \label{fig:policies.trade_off}
    \vspace{-0.2in}
  \end{center}
\end{figure}

Another possible method is to make the dropping decisions somewhere in between
source-only dropping and unrestricted dropping. From the graph in 
Figure~\ref{fig:policies.dataflow_graph}, we can see that if the dropping
decision is made at event {\tt C} then we can skip event {\tt E} with no wasted
work. If we instead choose to perform monitoring for event {\tt C}, then we want to perform all
monitoring on that flow. Similarly, event {\tt F} creates a decision point for
the bottom flow that will result in no wasted work and dropping event {\tt A}
allows the entire shown metadata flow to be dropped with no wasted work. If the
program can be
analyzed or profiled to identify these instructions which, if dropped, lead to
no wasted work, then this information can be used at run-time to minimize the
amount of wasted work. We refer to this policy as \emph{sub-flow dropping}.
Figure~\ref{fig:policies.min_drop} shows the points were dropping decisions are
made for this policy.
Min-event dropping is a coarser-grained decision than unrestricted dropping, but
finer-grained that source dropping. Thus, its ability to match overheads is
expected to sit between these other two policies. Similarly, there still exists
edge cases where wasted work can occur, but typically we expect much less
wasted work than unrestricted 
dropping. Thus, we expect the coverage achieved by split-point dropping for a
certain overhead to fall between source and unrestricted dropping. These three
dropping policies create a trade-off space between coverage achieved and
ability to match an overhead target.
Figure~\ref{fig:policies.trade_off} summarizes this trade-off between coverage
and ability to match overhead.
